---
title: "Statistical_Learning_Week5"
author: "Fatma Ozdal"
date: "2026-03-10"
output: pdf_document
---

```{r}
getwd()
setwd("~/Desktop")
```

```{r}
library(glmnet)
```

```{r}
#train <- readRDS(file.choose())
```

```{r}
#test <- readRDS(file.choose())
```

```{r}
train <- readRDS("masq_train.Rda")
test <- readRDS("masq_test.Rda")
```

```{r}
x_train <- model.matrix(D_DEPDYS ~ ., data = train)[, -1]
y_train <- train$D_DEPDYS

x_test <- model.matrix(D_DEPDYS ~ ., data = test)[, -1]
y_test <- test$D_DEPDYS
```

```{r}
set.seed(42) 

lasso_cv <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1)
ridge_cv <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0)
en_cv <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0.5)


min(lasso_cv$cvm); min(ridge_cv$cvm); min(en_cv$cvm)
```

```{r}
best_mod=lasso_cv 

eta_hat=predict(best_mod, newx = x_test, s = "lambda.min")
p_hat= 1 / (1 + exp(-eta_hat))


brier_score=mean((y_test - p_hat)^2)

predictions=ifelse(p_hat > 0.5, 1, 0)
mcr=mean(predictions != y_test)

cat("Brier Score:", brier_score, "\n")
cat("MCR:", mcr, "\n")
```

```{r}
best_coefs=coef(best_mod, s = "lambda.min")
selected_vars=row.names(best_coefs)[which(best_coefs != 0)]
print(selected_vars)
```
The results indicate most of variables belong to the Anhedonic Depression subscale. This shows that loss of pleasure is most useful symptom for identifying depressive disorder in this test.

```{r}
eta_hat=predict(lasso_cv, newx = x_test, s = "lambda.min")
p_hat =1 / (1 + exp(-eta_hat))

brier_score = mean((y_test - p_hat)^2)

predictions = ifelse(p_hat > 0.5, 1, 0)
mcr <- mean(predictions != y_test)

print(paste("Brier Score:", round(brier_score, 4)))
print(paste("Misclassification Rate (MCR):", round(mcr, 4)))
```


The MCR is 0.2264. It means that the model incorrectly predicted the diagnosis for about 22.6% of individuals in the test set. This also shows an overall accuracy of about 77.4%. 
The Brier score is 0.1614. 0.16 is a good result because lower score is better. 
All these results indicates that the penalized logistics regression model is a good predictor for this dataset.